---
title: "THmodel_output_commented"
author: "Sophia Ng, Josh Petrie, Richard Evans"
date: "`r format(Sys.time(),'%B, %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(magrittr))
```

Created by Josh Petrie on Jan 24, 2016, updated by Richard Evans on Oct 05, 2017

# Rationale

Using estimated parameter output from mcmc, this script creates output tables of:

1. mcmc acceptance rate, parameter means (median) and 95% CI
2. Serial interval mean (median) and 95% CI
3. Cumulative risks of community and household infection with 95% CI for all combinations of patient characteristics (e.g. <9 years, high risk, unvaccinated)

Each chain is an estimated parameter $\Theta = \left\{\theta_1,...,\theta_8\right\}$ from both the community hazard and HH hazard components of the model: $\Theta = \left\{\alpha, \gamma, \lambda_{HH}, \psi_C, \left\{\beta_i\right\}_{i=1}^n\right\}$

theta <- c(q, k, hh base haz, com haz scale, Age 9-17, Age 18+, HighRisk, Vaccination)

# Load Data

```{r}
setwd("S:/MartinEpi/Projects/BoV-Uganda/Graphs and Reports/HH-transmission-model-master_2/mcmc_output")
# saveRDS(model1, file = "model1.rds")
model1 <- readRDS("model1.rds")
setwd("S:/MartinEpi/Projects/BoV-Uganda/Graphs and Reports/HH-transmission-model-master_2/")
com <- read.csv("THmodel_exampleComData.csv") %>%
  filter(!is.na(COM))
```

# Acceptance Rate

optimal acceptance rate: ~45-55% in single dimension

```{r}
accept <- data.frame(
  theta = names(model1[1:8]),
  accept = round(colMeans(model1$accept, na.rm = TRUE), 2)
  )

a <- accept
row.names(a) <- c("alpha", "gamma", "HH base haz", "Com haz scale", 
                  "Age 9-17", "Age 18+", "HighRisk", "Vaccination")
a
```

# Posterior means, medians, CIs

```{r}
ci.fn <- function(x) quantile(x, c(0.025, 0.975))

mu <- round(sapply(model1[1:8], mean), 6)
median <- round(sapply(model1[1:8], median), 6) # very similar

ci <- round(sapply(model1[1:8], ci.fn), 6) %>%
  t() %>%
  data.frame() %>%
  rename(LB = X2.5., UB = X97.5.)

param.output <- data.frame(accept, mu, ci) %>%
  mutate(theta = factor(theta, labels = row.names(a)))
row.names(param.output) <- c()

param.output
```

# Calculate Serial Intervals

Calculate serial interval based on Weibull distribution, must provide $\alpha, \gamma$ (shape and scale) parameters from dataset

```{r}
SI.FN <- function(params, HHt, shape, scale) {
  SI.FN <- exp(-(HHt / params[[shape]]) ^ params[[scale]]) -
    exp(-((HHt + 1) / params[[shape]]) ^ params[[scale]])
  SI.FN
}

SI.fn <- function(params, HHt, shape, scale) {
  SI.instance <- SI.FN(params, HHt, shape, scale)
  SI <- sum(HHt * SI.instance) / sum(SI.instance)
  SI
}

SI.apply <- function(params, HHt, shape, scale) {
  SIvec <- vector(length = length(params[[shape]]))
  temp <- data.frame(
    theta1 = params[[shape]],
    theta2 = params[[scale]]
  )
  for (i in 1:length(temp[, 1])) 
    SIvec[i] <- SI.fn(temp[i, ], HHt, shape, scale)
  SIvec
}
```

Mean Serial Interval

```{r}
muSI <- SI.apply(model1,
                 seq(0, 14, 1),
                 shape = "theta1", 
                 scale = "theta2") %>%
  mean() %>%
  round(6)

ciSI <- SI.apply(model1, 
                 seq(0, 14, 1), 
                 shape = "theta1", 
                 scale = "theta2") %>%
  quantile(c(0.025, 0.975)) %>%
  round(6)

# save output
SI.output <- data.frame(SI = muSI, ciSILwr = ciSI[1], ciSIupr = ciSI[2])
row.names(SI.output) <- c()
SI.output
```

# Calculate total HH risk of infection

```{r}
# HHRISK calculates the total household risk of infection
#HHRISK calculates the total household risk of infection from a given set of parameters
HHRISK <- function(params, HHt, AGE9to17, AGEgte18, HighRisk, Vax) {
  #### why is this here?
  SI.FN <- exp(-(HHt / params$theta1) ^ params$theta2) - 
    exp(-((HHt + 1) / params$theta1) ^ params$theta2)
  
  lin.pred <- params$theta5 * AGE9to17 + 
    params$theta6 * AGEgte18 + 
    params$theta7 * HighRisk + 
    params$theta8 * Vax
  HH.HAZ <- SI.FN * params$theta3 * exp(lin.pred)
  eHH.HAZ <- exp(-HH.HAZ)
  
  HHl <- loess(1 - eHH.HAZ ~ HHt, span = .35)
  HHf <- function(x) predict(HHl, newdata = x)
  HHi <- integrate(HHf, 0, 14)
  
  HHi$value
}

#HHRISK.apply applies HHRISK to parameters generated from each mcmc iteration
HHRISK.apply <- function(params, HHt, AGE9to17, AGEgte18, HighRisk, Vax) {
  HouseRisk <- vector(length = length(params[[1]]))
  temp <- data.frame(do.call(cbind, params[1:8]))
  for(i in 1:nrow(temp)) {
    HouseRisk[i] <- HHRISK(temp[i,], HHt, AGE9to17, AGEgte18, HighRisk, Vax)
  } 
  HouseRisk
}
```

calculate mean household risk of infection for each combination of subject characteristics

```{r}
muHH <- c(mean(HHRISK.apply(model1, seq(0, 14, 1), 0, 0, 0, 0)),
          mean(HHRISK.apply(model1, seq(0, 14, 1), 0, 0, 1, 0)),
          mean(HHRISK.apply(model1, seq(0, 14, 1), 0, 0, 0, 1)),
          mean(HHRISK.apply(model1, seq(0, 14, 1), 0, 0, 1, 1)),
          mean(HHRISK.apply(model1, seq(0, 14, 1), 1, 0, 0, 0)),
          mean(HHRISK.apply(model1, seq(0, 14, 1), 1, 0, 1, 0)),
          mean(HHRISK.apply(model1, seq(0, 14, 1), 1, 0, 0, 1)),
          mean(HHRISK.apply(model1, seq(0, 14, 1), 1, 0, 1, 1)),
          mean(HHRISK.apply(model1, seq(0, 14, 1), 0, 1, 0, 0)),
          mean(HHRISK.apply(model1, seq(0, 14, 1), 0, 1, 1, 0)),
          mean(HHRISK.apply(model1, seq(0, 14, 1), 0, 1, 0, 1)),
          mean(HHRISK.apply(model1, seq(0, 14, 1), 0, 1, 1, 1)))
round(muHH, 6)
```
 
calculate 95% CI for household risk of infection for each combination of subject characteristics

```{r}
f <- function(func.output) quantile(func.output, c(0.025, 0.975))
ciHH <- c(f(HHRISK.apply(model1, seq(0, 14, 1), 0, 0, 0, 0)),
          f(HHRISK.apply(model1, seq(0, 14, 1), 0, 0, 1, 0)),
          f(HHRISK.apply(model1, seq(0, 14, 1), 0, 0, 0, 1)),
          f(HHRISK.apply(model1, seq(0, 14, 1), 0, 0, 1, 1)),
          f(HHRISK.apply(model1, seq(0, 14, 1), 1, 0, 0, 0)),
          f(HHRISK.apply(model1, seq(0, 14, 1), 1, 0, 1, 0)),
          f(HHRISK.apply(model1, seq(0, 14, 1), 1, 0, 0, 1)),
          f(HHRISK.apply(model1, seq(0, 14, 1), 1, 0, 1, 1)),
          f(HHRISK.apply(model1, seq(0, 14, 1), 0, 1, 0, 0)),
          f(HHRISK.apply(model1, seq(0, 14, 1), 0, 1, 1, 0)),
          f(HHRISK.apply(model1, seq(0, 14, 1), 0, 1, 0, 1)),
          f(HHRISK.apply(model1, seq(0, 14, 1), 0, 1, 1, 1)))
round(ciHH, 6)
```

Calculate total community risk of infection 

```{r}
#CRISK calculates the total community risk of infection
CRISK <- function(params, com, Ct, AGE9to17, AGEgte18, HighRisk, Vax) {
  lin.pred <- params$theta5 * AGE9to17 + 
              params$theta6 * AGEgte18 + 
              params$theta7 * HighRisk + 
              params$theta8 * Vax
  C.HAZ <- com$COM * params$theta4 * exp(lin.pred)
  eC.HAZ <- exp(-C.HAZ)
  
  Cl <- loess(1 - eC.HAZ ~ Ct, span = .25)
  Cf <- function(x) predict(Cl, newdata = x)
  Ci <- integrate(Cf, 0, 104)
  Ci$value
}

#CRISK.apply applies CRISK to parameters generated from each mcmc iteration
CRISK.apply <- function(params, com, Ct, AGE9to17, AGEgte18, HighRisk, Vax) {
  ComRisk <- vector(length = length(params[[1]]))
  temp <- data.frame(do.call(cbind, params[1:8]))
  for(i in 1:nrow(temp)){
    ComRisk[i] <- CRISK(temp[i, ], com, Ct, AGE9to17, AGEgte18, HighRisk, Vax)
  }
  ComRisk
}
```

calculate mean community risk of infection for each combination of subject characteristics

```{r}
muC <- c(mean(CRISK.apply(model1, com, seq(0, 104, 1), 0, 0, 0, 0)),
         mean(CRISK.apply(model1, com, seq(0, 104, 1), 0, 0, 1, 0)),
         mean(CRISK.apply(model1, com, seq(0, 104, 1), 0, 0, 0, 1)),
         mean(CRISK.apply(model1, com, seq(0, 104, 1), 0, 0, 1, 1)),
         mean(CRISK.apply(model1, com, seq(0, 104, 1), 1, 0, 0, 0)),
         mean(CRISK.apply(model1, com, seq(0, 104, 1), 1, 0, 1, 0)),
         mean(CRISK.apply(model1, com, seq(0, 104, 1), 1, 0, 0, 1)),
         mean(CRISK.apply(model1, com, seq(0, 104, 1), 1, 0, 1, 1)),
         mean(CRISK.apply(model1, com, seq(0, 104, 1), 0, 1, 0, 0)),
         mean(CRISK.apply(model1, com, seq(0, 104, 1), 0, 1, 1, 0)),
         mean(CRISK.apply(model1, com, seq(0, 104, 1), 0, 1, 0, 1)),
         mean(CRISK.apply(model1, com, seq(0, 104, 1), 0, 1, 1, 1)))
round(muC, 6)
```

calculate 95% CI for community risk of infection for each combination of subject characteristics

```{r}
f <- function(func.output) quantile(func.output, c(0.025, 0.975))
ciC <- c(f(CRISK.apply(model1, com, seq(0, 104, 1), 0, 0, 0, 0)),
         f(CRISK.apply(model1, com, seq(0, 104, 1), 0, 0, 1, 0)),
         f(CRISK.apply(model1, com, seq(0, 104, 1), 0, 0, 0, 1)),
         f(CRISK.apply(model1, com, seq(0, 104, 1), 0, 0, 1, 1)),
         f(CRISK.apply(model1, com, seq(0, 104, 1), 1, 0, 0, 0)),
         f(CRISK.apply(model1, com, seq(0, 104, 1), 1, 0, 1, 0)),
         f(CRISK.apply(model1, com, seq(0, 104, 1), 1, 0, 0, 1)),
         f(CRISK.apply(model1, com, seq(0, 104, 1), 1, 0, 1, 1)),
         f(CRISK.apply(model1, com, seq(0, 104, 1), 0, 1, 0, 0)),
         f(CRISK.apply(model1, com, seq(0, 104, 1), 0, 1, 1, 0)),
         f(CRISK.apply(model1, com, seq(0, 104, 1), 0, 1, 0, 1)),
         f(CRISK.apply(model1, com, seq(0, 104, 1), 0, 1, 1, 1)))
round(ciC, 6)
```

```{r}
labels <- c("<9,   HighRisk -, Vax -", 
            "<9,   HighRisk +, Vax -", 
            "<9,   HighRisk -, Vax +", 
            "<9,   HighRisk +, Vax +", 
            "9-17, HighRisk -, Vax -", 
            "9-17, HighRisk +, Vax -", 
            "9-17, HighRisk -, Vax +",
            "9-17, HighRisk +, Vax +", 
            "18+,  HighRisk -, Vax -", 
            "18+,  HighRisk +, Vax -", 
            "18+,  HighRisk -, Vax +", 
            "18+,  HighRisk +, Vax +")

#put mean and 95% community and household risks in a data frame
risk.output <- data.frame(
  labels = labels, 
  HHRisk = muHH, 
  ciHHlwr = ciHH[c(1,3,5,7,9,11,13,15,17,19,21,23)], 
  ciHHupr = ciHH[c(2,4,6,8,10,12,14,16,18,20,22,24)],
  CRisk = muC,
  ciClwr = ciC[c(1,3,5,7,9,11,13,15,17,19,21,23)],
  ciCupr = ciC[c(2,4,6,8,10,12,14,16,18,20,22,24)]
  )

risk.output
```

plot posterior distribution chains

```{r, fig.height = 16, fig.width = 16}
chains <- do.call(cbind, model1[1:8]) %>%
  data.frame()

par(mfrow = c(2, 4))
# chain trace plots
for (i in 1:ncol(chains)) 
  plot(chains[, i], type = "l",
       ylab = paste("theta", i, sep = ""),
       xlab = 'MCMC iteration')

# density plots
for (i in 1:ncol(chains))
  plot(density(chains[, i]),
       main = paste("theta", i, sep = ""))

# histograms
for (i in 1:ncol(chains))
  hist(chains[, i], 
       main = paste("theta", i, sep = ""),
       xlab = '')

# ACF plots
for (i in 1:ncol(chains)) acf(chains[, i])

# Partial ACF plots
for (i in 1:ncol(chains)) pacf(chains[, i])

# coefficient/hazard estimate plot with credible intervals
param.output %>%
  ggplot(aes(x = theta, y = mu)) %>%
  add(coord_flip()) %>%
  add(geom_point()) %>%
  add(geom_linerange(aes(ymin = LB, ymax = UB))) %>%
  add(theme_minimal(16)) %>%
  add(labs(
    y = expression(mu ~ '(95% CI)'),
    x = ""
  ))

GGally::ggpairs(chains)
```

